US President Prediction

### 2024-11-02 paper/paper.qmd
update survey questions

As indicated in the @sec-data, we do not have a confident belief of priors of parameters of our model since Harris replace the role of Biden by indecent and we do not have enough data about her previous polling outcomes as a president candidate compared to Biden. Therefore, we would not use the bayesian model but instead using a more simple multilinear regression model which predicts the support rate of Trump and Harris using time towards elections date and pollsters as they are beliveved to highly corelated to the response variable in @sec-data. We would bulid 8 models whihc represent national trend and that of seven swing states. 
## Model set-up

Define $y_{i,\text{stateA}}$ as the percentage of support for a candicate in a poll in the specific state.

$$
y_{i,national} = \beta_{0,national} + \beta_{1,time,national} + \beta_{2,pollster,national} + \epsilon_{national}
$$ 
$$
y_{i,\text{AZ}} = \beta_{0,\text{AZ}} + \beta_{1,\text{time,AZ}} + \beta_{2,\text{pollster,AZ}} + \epsilon_{\text{AZ}}
$$
$$
y_{i,\text{FL}} = \beta_{0,\text{FL}} + \beta_{1,\text{time,FL}} + \beta_{2,\text{pollster,FL}} + \epsilon_{\text{FL}}
$$
$$
y_{i,\text{GA}} = \beta_{0,\text{GA}} + \beta_{1,\text{time,GA}} + \beta_{2,\text{pollster,GA}} + \epsilon_{\text{GA}}
$$
$$
y_{i,\text{MI}} = \beta_{0,\text{MI}} + \beta_{1,\text{time,MI}} + \beta_{2,\text{pollster,MI}} + \epsilon_{\text{MI}}
$$
$$
y_{i,\text{NC}} = \beta_{0,\text{NC}} + \beta_{1,\text{time,NC}} + \beta_{2,\text{pollster,NC}} + \epsilon_{\text{NC}}
$$
$$
y_{i,\text{PA}} = \beta_{0,\text{PA}} + \beta_{1,\text{time,PA}} + \beta_{2,\text{pollster,PA}} + \epsilon_{\text{PA}}
$$
$$
y_{i,\text{WI}} = \beta_{0,\text{WI}} + \beta_{1,\text{time,WI}} + \beta_{2,\text{pollster,WI}} + \epsilon_{\text{WI}}
$$

time:The date when the pollster starts polling

pollster: The pollster which run the polling

@@

# Results {#set-result}

@@

7.

8.

9.

>>>>

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix @sec-model-details].

## Model set-up

Define $y_i$ as the percentage of support for a candicate in a poll.

${y_i} = \beta_0 + \beta_1 pollscore + \beta_2 methodology + \beta_3transparency score + \beta_4 hypothetical + \epsilon$

pollscore: A numeric value representing the score or reliability of the pollster in question (e.g., -1.1). "The error and bias we can attribute to a pollster. Negative numbers are better. Stands for "Predictive Optimization of Latent skill Level in Surveys, Considering Overall Record, Empirically.

methodology: The method used to conduct the poll (e.g., Online Panel).

transparency_score: A score reflecting the pollster’s transparency about their methodology (e.g., 9.0). "A grade for how transparent a pollster is, calculated based on how much information it discloses about its polls and weighted by 
by recency. The highest Transparency Score is 10."

hypothetical: Indicates whether the poll is about a hypothetical match-up.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Library setup
#install.packages("glm")
library(glmnet)
library(ggplot2)
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-model1
#| tbl-cap: "model 1"
#| warning: false
# Get the analysis data
analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

# Model
glm_model1 <- glm(pct ~ pollscore + methodology + transparency_score + hypothetical, data = analysis_data, family = gaussian())

# Get the summary
model_summary <- summary(glm_model1)

# Convert the summary coefficients to a data frame
summary_table <- as.data.frame(model_summary$coefficients)

# Create a table with a caption and control the size
kable(summary_table, caption = "Model 1 Summary Table", size = "small")
```

## Lasso Regularization {#sec-model-lassoregularization}
Lasso Regularization is performed to see if the selected variable in our model would best predict the results of polls percentage of a candidate. If the variable do not align with our actual numbers, we should consider other variables or a interaction variable.

```{r}
#| echo: false
#| eval: true
#| label: tbl-lassoregularization
#| tbl-cap: "lasso regularization"
#| warning: false
# Prepare the predictors
x <- model.matrix(pct ~ pollscore + methodology + transparency_score + hypothetical, data = analysis_data)[, -1]

# Prepare the response variable
y <- analysis_data$pct

# Lasso regression (alpha = 1)
lasso_model <- glmnet(x, y, alpha = 1, family = "gaussian")

# Cross-validation for Lasso
cv_lasso <- cv.glmnet(x, y, alpha = 1, family = "gaussian")

plot(cv_lasso)

# Get the best lambda
best_lambda_lasso <- cv_lasso$lambda.min

# Refit Lasso with the optimal lambda
lasso_final_model <- glmnet(x, y, alpha = 1, lambda = best_lambda_lasso, family = "gaussian")

# View Lasso coefficients
coef.lasso <- coef(lasso_final_model)

# Predict using Lasso
pred_lasso <- predict(lasso_final_model, s = best_lambda_lasso, newx = x)

```

Computing the Mean Square Error would give us a sense of our lasso model prediction. 
```{r}
#| echo: false
#| eval: true
#| label: tbl-msr
#| tbl-cap: "Mean Squared Error"
#| warning: false
# Calculate Mean Squared Error
mse <- mean((y - pred_lasso)^2)
cat("Mean Squared Error:", mse, "\n")

# Calculate R-squared
sst <- sum((y - mean(y))^2)
sse <- sum((y - pred_lasso)^2)
r_squared <- 1 - (sse / sst)
cat("R-squared:", r_squared, "\n")

# Create a data frame for plotting
results_df <- data.frame(
  Metric = c("SST", "SSE", "R-squared"),
  Value = c(sst, sse, r_squared)
)

kable(results_df, caption = "Mean Squared Error Table")
```

From this graph, we can see that the cluster points do not perfectly align on the line of the best fit so we should still investigate in other variables. 
```{r}
#| echo: false
#| eval: true
#| label: tbl-actualvspredicted
#| tbl-cap: "actual vs predicted"
#| warning: false

# Compare predicted vs actual values
comparison <- data.frame(Actual = y, Predicted = as.vector(pred_lasso))

kable(head(comparison), caption = "Actual vs Predicted")

# Plot actual vs predicted values
ggplot(comparison, aes(x = Actual, y = Predicted)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  labs(title = "Actual vs Predicted Values",
       x = "Actual Values",
       y = "Predicted Values") +
  theme_minimal()
```

@@

7. sex

8. race

9. education
